### 2011-05-30 thesis/novelty-intro.tex
More novelty detection introduction and history text from the tech report.

This chapter presents a statis

Laying down base concepts that are then used in \autoref{chap:novelty} to
implement novelty detection on the semantic spatial representation.

It start by presenting a brief overview on novelty detection.

It introduces novelty as a

Shows novelty has an associated ordering.
Shows novelty has a threshold function.

@@

\begin{itemize}
\item Introduction
\item History
\item Approaches to Novelty Detection
\end{itemize}

\section{Novelty Detection Review and Related Work}
% Introduce novelty detection
Novelty detection deals with detecting that a new input was generated by a class
other than the ones the system knows about~\cite{markou2003novelty}.
It is harder than classification as only positive samples of a class are available
rendering normal classification methods invalid.

@@

\subsection{Unconditional Probability}
The unconditional probability $P(x)$ plays an important role on obtaining a
correct order relation for performing novelty detection.
It serves as a normalizing component that allows the system to  whether
a given sample conditional probability arises from it belonging to the
known concept or from it likelihood of being sampled.

@@

%\subsection{Novelty Detection on a Variable Set of Features}
Note that very often, novelty detection is applied on a fixed set of features
together with an assumption of a uniform unconditional probability.
On those cases $P(x)$ becomes a constant and therefore a novelty threshold
can be directly applied on $P(x|\overline{novel})$ as is the case in \cite{bishop1994novelty}.
But in the case where the set of features $x$ is variable it cannot be
discarded. There $P(x)$ plays a role in levering all the conditional
probabilities on different sets of variables into the same measure units
such that a threshold can be implemented.

>>>>

This chapter presents concepts that are used as base when adding novelty
detection capabilities on the semantic representation in \autoref{chap:novelty}.
It starts by providing the reader with a brief overview on novelty detection
techniques and related works.

After it shows that an optimal novelty detection system can be implemented by
thresholding on a order-relation defined over the inputs.
And that an equivalent order-relation can be imposed by a ratio between
conditional and unconditional probabilities.
Some discussion on how to interpret the meaning of those factors is also
presented.

@@

\section{Novelty Detection Review and Related Work}
% Introduce novelty detection
%
%
%

Novelty detection, also called outlier or anomaly detection, is the
classification problem related with identification of new or unknown data that
the system was not aware of during the training.

It has several applications such as fault detection, intrusion detection,
detection of masses in mammograms, hand written digit recognition and many
others~\citep{markou2003novelty}.
And has seen an increase of its application in the area of robotics and control.

It is posed as a one class classification where there is only access to the
positive samples. The objective then is to classify new inputs as being
generated by the class it was trained or if they belong to some new unknown one.

Either by impossibility of generating samples of "unknown", or by extreme
difficulty to acquire negative-examples, two-class classifiers cannot be
applied to this problem.
Several other methods were developed to address this problem and they often can
be categorized in the following approaches:
based on probability density estimation;
based on using the feature space to separate the given class from others, such as one-class \gls{SVM};
and based on reconstruction-error from a lower-dimension.

\subsubsection*{Density estimation}
One of the simplest approach for novelty detection is pose the problem as a
probability density estimation problem and use the expected probability of a
given input to trigger the case as novel.

Examples of such approaches are Gaussian Mixtures Models and Parzen-window estimators.
In order to reduce the input size, those approaches often employ
dimension-reduction techniques such as \gls{PCA}.

This approach is used in \cite{bishop1994novelty} where \emph{novelty detection}
is added to a \emph{neural network} system. On it the density estimation on the
input is calculated using a \emph{Parzen window} over training-data and that
value is used as a novelty measure. This way they detect novel data that their
network has not seen before and so the expected output of their system is not
reliable.

\subsubsection*{One-class {SVM}}
One-class \gls{SVM} approaches try to distinguish the class by separating the
training set from all other points in the input space. They try to archive that
by enclosing the training set by some structure (example: hyper-sphere)~\citep{scholkopf2000support}.
This leads to a decision boundary and follows the same principle as Vapnik
solving classification with \gls{SVM}: do not solve something harder.

\subsubsection*{Reconstruction Error}
Error reconstruction methods for novelty detection use the assumption that the
class to be defined lies on a manifold embedded on the input space of higher
dimensions.
By using dimensional reduction techniques, they try to defined that manifold
and calculate how far a new input is from it.

\label{sec:kernel-pca}
One of the most common methods used for that is
\gls{K-PCA}~\citep{scholkopf1997kernel}, which uses the kernel-trick
(\autoref{sec:kernel-trick}) to extend \gls{PCA} and perform a nonlinear
dimensionality reduction on the input. This technique has been successfully
used in novelty detection by \cite{Hoffmann2007863}.

Novelty detection deals with detecting that a new input was generated by a class
other than the ones the system knows about~\cite{markou2003novelty}.
It is harder than classification as only positive samples of a class are available
rendering normal classification methods invalid.

@@

\subsection{Unconditional Probability}
The unconditional probability $P(x)$ plays an important role on obtaining a
correct order relation for performing novelty detection.
It serves as a normalizing component that allows the system to figure out
whether a given sample conditional probability arises from it belonging to
the known concept or from it likelihood of being sampled.

@@

%\subsection{Novelty Detection on a Variable Set of Features}
Note also that often novelty detection is applied on a fixed set of features
together with an assumption of a uniform unconditional probability.
On those cases $P(x)$ becomes a constant and therefore a novelty threshold
can be directly applied on $P(x|\overline{novel})$ as is the case in \cite{bishop1994novelty}.
But in the case where the set of features $x$ is variable it cannot be
discarded. There $P(x)$ plays a role in levering all the conditional
probabilities on different sets of variables into the same measure units
such that a threshold can be implemented.
